\textbf{ Accurate spectroscopic estimates of luminosity and distance\\ for luminous red-giant stars }

### 2018-07-25 paper/sp.tex
outline, acks, words

% To-do items
% -----------
% - draft introduction
% - draft data+results section
% - draft discussion
% - write list of things we never do!
% - make sure we address the cross-validation point
% - make sure we address the sensibleness and generality of the linear model for the photometric space

@@

\author{Melissa Ness}
\affiliation{Flatiron Institute}
\affiliation{Department of Astronomy, Columbia University}

@@

HOGG: LIST of things we don't do, like don't cut the sample on
astrometric quantities, and why not!! And like don't assume distance
is inverse-measured-parallax.

@@

Under the assumptions stated in the previous Section, there are not many choices
for building a model of the stars that is justifiable probabilistically.
Here we lay out the model and methodology, and apply it in the following Section.

Since the photometry is (by the ``good features'' assumption above)
better measured than the parallaxes, we
presume that the parallax uncertainties dominate
the uncertainties in the luminosity measurements.
For each star $n$ in our training set, we define an inverse-square-root
luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
\begin{eqnarray}
Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
\\
\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
\quad,
\end{eqnarray}
where
$\varpi_n$ is the \gaia\ Catalog entry for the parallax for star $n$,
$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
for star $n$.
In what follows, the $Q_n$ are the training data.
Naively, the $Q_n$ are related to the absolute magnituges $M_{Kn}$ for the stars by
\begin{eqnarray}
M_{Kn} &\approx& 5\,\log_{10} Q_n
\quad.
\end{eqnarray}
These would be the absolute magnitudes of the stars under the
``the distance is the inverse of the
measured parallax'' assumption, although we never, in this work, ever make that
assumption, by God.
We construct the $Q_n$ because they are related directly to the absolute magnitudes,
but they (under assumptions) are subject to known Gaussian noise.

@@

The \gaia\ Catalog describes a likelihood function!

There is a parallax offset.

We are only looking at \apogee\ luminous giants (selected spectroscopically).

We use photometry from \gaia, \zmass, \wise.

@@

@@

@@

\acknowledgements
This project was developed in part at the
2018 \acronym{NYC} Gaia Sprint, hosted by the Center for Computational Astrophysics of
the Flatiron Institute in New York City.

>>>>

% To-do items
% -----------
% - draft introduction
% - draft data+results section
% - draft discussion
% - make sure we address the cross-validation point
% - make sure we address the sensibleness and generality of the linear model for the photometric space
% - address convexity of the model...?

@@

@@

In addition to making the above assumptions, we also avoid various practices
with the data that are tempting but lead to strong biases.
For example, we never cut on \gaia\ parallax or parallax signal-to-noise.
The common practices of cutting to parallaxes that are good to 20~percent,
or Catalog parallaxes that are positive, or parallaxes that are smaller or
larger than something, are all practices that will bias results on stellar
collections.
That is, if you cut on parallax or parallax signal-to-noise and you subsequently
take an average of parallaxes for some population, or perform a regression (as we
do here), the results of that average or regression will be strongly biased.
We never cut on parallax or parallax signal-to-noise.
By using a justifiable likelihood function (the ``likelihood'' assumption above),
we can use all of the \gaia\ parallaxes without the low signal-to-noise and
negative parallaxes causing trouble for our regressions, and without the biases
that enter when cuts are made.

Along similar lines, we never assume that the distance is the inverse of the
measured parallax.
In what follows, a star's distance is a latent property of the star, which generates
the \gaia\ Catalog parallax through a noisy process (again, this is
the ``likelihood'' assumption above).
We never take the inverse or the logarithm of the measured parallax at any time.
This is related to the no-cuts point above:
If you take the inverse or logarithm of the parallax, you can't operate safely
on the negative and low signal-to-noise parallaxes, which in turn will require
making cuts on the sample, which will in turn bias the results.

Finally, we never use Lutz--Kelker corrections (HOGG CITE) or distance
posteriors (HOGG CITE). These both involve (implicit or explicit) priors on 
distance.
When multiple stars are combined as we combine them here (below),
use of distance posteriors instead of parallax likelihoods is not just
unjustified, but it also leads to an
effective raising of the distance prior to an enormous power.
That is, the effective prior on a $N$-star inference performed naively
with distance posteriors (made with a weak prior) can end up
bringing into the inference an \emph{exceedingly
strong prior}.
Therefore we don't use any prior-contaminated parallax or distance
inputs to the inference.

@@

There are not many choices
for building a model of the stars that is both justifiable probabilistically
and consistent with the assumptions stated in the previous Section.
Here we lay out the model and methodology.
We apply the method to real data in the following Sections.

Since the photometry is (by the ``good features'' assumption above)
better measured than the parallaxes, we
presume that the parallax uncertainties dominate
the uncertainties in the luminosity measurements.
For each star $n$ in our training set, we define an inverse-square-root
luminosity $Q_n$ estimate that will be subject to Gaussian noise under these assumptions.
\begin{eqnarray}
Q_n &\equiv& \varpi_n\,10^{0.2\,(K_n - 10)}
\\
\sigma_n &\equiv& \sigma_{\varpi n}\,10^{0.2\,(K_n - 10)}
\quad,
\end{eqnarray}
where
$\varpi_n$ is the \gaia\ Catalog entry for the parallax (in mas) for star $n$,
$K_n$ is the \zmass\ Catalog $K_s$-band magnitude for star $n$,
the $Q_n$ have Gaussian errors of root-variance $\sigma_n$,
and $\sigma_{\varpi n}$ is the \gaia\ Catalog entry for the uncertainty on parallax
for star $n$.
In what follows, the $Q_n$ are the training data.
Naively, the $Q_n$ are related to the absolute magnitudes $M_{Kn}$ for the stars by
\begin{eqnarray}
M_{Kn} &\approx& 5\,\log_{10} Q_n
\quad.
\end{eqnarray}
These would be the absolute magnitudes of the stars under the
``the distance is the inverse of the
measured parallax'' assumption, although we never, in this work, ever make that
assumption, by God.
We construct the $Q_n$ because they are related directly to the absolute magnitudes,
but they (under assumptions) are subject to known Gaussian noise.

@@

We take the \apogee\ data from what subset?

We cut on $\logg$ how and why? Comment on the trustworthiness of the $\logg$ values.

We take the \gaia\ data and apply the parallax offset.

We match \apogee\ to \gaia\ how?

We match the catalog to \zmass\ and \wise\ how?

Figures that show off the data.

@@

We construct the feature vectors $x_n$ how and what is their dimensionality $D$?

Brief comment on the flexibility of the photometric model.

We do 2-fold train-and-test to make results independent of \gaia.

We set the regularization parameters how? And why?

We compute our precision how? Figures.

@@

(In addition: what did we assume about the \apogee\ data: Good $\logg$ values.

@@

\acknowledgements
It is a pleasure to thank
  Daniel Michalik (Lund), and
  Melissa Ness (Columbia)
for help with this project.
This project was developed in part at the
2018 \acronym{NYC} Gaia Sprint, hosted by the Center for Computational Astrophysics of
the Flatiron Institute in New York City.
